---
title: "Prices of houses"
author: "Shuaib Akinyemi"
date: "`r Sys.Date()`"
output:
  slidy_presentation: default
  powerpoint_presentation: default
  beamer_presentation: default
  ioslides_presentation: default
---

## Background
Nigerian property prices are notorious for fluctuating over time. This can be explained by several causes, including the naira's decline in value. However, this investigation aims to determine how property values alter depending on location, number of rooms, and other intrinsic elements. This study will also assist us in determining how a house’s inherent features, such as bedrooms and parking, affect its pricing.


## Preparation Of Data and Required Packages
The dataset to be used was extracted from data.world website. We will be loading the tidyverse package which is required for this analysis and also load the data using the read.csv function.
     

```{r}
library(tidyverse)
nigeria_houses <- read.csv("~/nigeria_houses_data.csv")
     
```

#view the data
```{r}
glimpse(nigeria_houses)
head(nigeria_houses)
```
The data is clean and processed so that means it can be relied on for our analysis.


Analysis
Average price for 3 bedroom flats across different states in Nigeria
```{r}
nigeria_houses %>%
  drop_na() %>%
  filter(bedrooms == 3) %>%
  group_by(state) %>%
  summarize(avg_price = mean(price)) %>% 
  arrange(-avg_price)
```
Towns with the cheapest price for 3 bedroom flats in Lagos
```{r}
nigeria_houses %>%
  drop_na() %>%
  filter(bedrooms == 3 & state == "Lagos") %>%
  group_by(town) %>%
  summarize(avg_price = mean(price)) %>% 
  arrange(avg_price)
```

lowest to highest prices for 3 bedroom flats in Abuja
```{r}
nigeria_houses %>%
  drop_na() %>%
  filter(bedrooms == 3 & state == "Abuja") %>%
  group_by(town) %>%
  summarize(avg_price = mean(price)) %>% 
  arrange(avg_price)
```

highest to lowest prices for 3 bedroom flats in Abuja

```{r}
nigeria_houses %>%
  drop_na() %>%
  filter(bedrooms == 3 & state == "Abuja") %>%
  group_by(town) %>%
  summarize(avg_price = mean(price)) %>% 
  arrange(-avg_price)
```
Prices of 3 bedroom flats according to types or title

```{r}
nigeria_houses %>%
  drop_na() %>%
  filter(bedrooms == 3) %>%
  group_by(title) %>%
  summarize(avg_price = mean(price)) %>% 
  arrange(-avg_price)
```
Visualization
We will be using visualizations to show how the price of houses is affected by the different inherent characteristics such as bedroom and parking space.

Relationship Between Number Of Toilet and Average Price Of Houses.
```{r}
nigeria_houses %>% 
  drop_na() %>% 
  filter(bedrooms == 3) %>% 
  group_by(toilets) %>% 
  summarize(avg_price = mean(price)) %>% 
  ggplot(mapping = aes(x = toilets, y = avg_price, color = toilets)) +
  geom_line() +
  labs(title = 'Relationship Between Number Of Toilet and Average Price Of Houses', subtitle = 'Emphasis On Three Bedroom Flats') +
  scale_color_continuous(name = "Number of Toilets")


     
```
There isn't much correlation going on there.



Relationship Between Number Of Bathrooms and Average Price Of Houses
```{r}
nigeria_houses %>% 
  drop_na() %>% 
  filter(state == 'Lagos' & town == 'Victoria Island (VI)' & title == 'Terraced Bungalow') %>% 
  group_by(bathrooms) %>% 
  summarize(avg_price = mean(price)) %>% 
  ggplot(mapping = aes(x = bathrooms, y = avg_price, color = "green")) +
  geom_point() +
  labs(title = 'Relationship Between Number Of Bathrooms and Average Price Of Houses', subtitle = 'Emphasis On Terraced Bungalows in Victoria Island') +
  scale_color_manual(values = "green")

```
The number of bathrooms also seems not to be a determining factor for the prices of houses.


Relationship Between Number Of Parking Space and Average Price Of Houses
```{r}
nigeria_houses %>% 
  drop_na() %>% 
  filter(state == 'Lagos' & town == 'Victoria Island (VI)' & title == 'Terraced Bungalow') %>% 
  group_by(parking_space) %>% 
  summarize(avg_price = mean(price)) %>% 
  ggplot(mapping = aes(x = parking_space, y = avg_price, fill = "red")) +
  geom_bar(stat = "identity") +
  labs(title = 'Relationship Between Number Of Parking Space and Average Price Of Houses', subtitle = 'Emphasis On Terraced Bungalows in Victoria Island')


```
This shows some direct correlation.

Relationship Between Number Of Bedrooms and Average Price Of Houses
```{r}
nigeria_houses %>% 
  drop_na() %>% 
  filter(state == 'Lagos' & town == 'Victoria Island (VI)' & title == 'Terraced Bungalow') %>% 
  group_by(bedrooms) %>% 
  summarize(avg_price = mean(price)) %>% 
  ggplot(mapping = aes(x = bedrooms, y = avg_price)) +
  geom_line() +
  labs(title = 'Relationship Between Number Of Bedrooms and Average Price Of Houses', subtitle = 'Emphasis On Terraced Bungalows in Victoria Island')

```
Conclusion
The prices of houses are not really determined by the number of bathrooms or toilets but by the number of bedrooms and parking space available.
The five most expensive states to buy a house in Nigeria are Lagos, Abia, Abuja, Delta and Anambra.
The most expensive place to buy a house in Lagos is Ikoyi while the cheapest place is Imota.
Gwagwalada is the cheapest place to buy a house in Abuja and the most expensive houses are in Maitama.
The most expensive houses in Nigeria are semi detached duplex while the cheapest are semi detached bungalows.

